### 2021-12-01 paper/output/document/application.tex
added will notes

%
\begin{isabellebody}%
\setisabellecontext{application}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Applications \label{sec:applications}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In this chapter, I 
demonstrate my system's ability to formalize longer, more complicated ethical arguments and understand 
the additional capabilities necessary to use my system in practice. In Chapter 2
and Chapter 3, I performed metaethical reasoning while testing different formalizations of the FUL. Metaethical
reasoning analyzes properties of moral thought itself and involves questions about the nature of ethical
truth. This kind of reasoning contrasts with ``applied 
ethical reasoning," which is the use of ethics to resolve dilemmas and make judgements about 
what an agent should or should not do. Applied ethical reasoning is relevant with respect to an agent's
particular situation, not just an an ethical theory as an abstract entity. In previous chapters' application tests,
I performed some toy examples of applied ethical reasoning, but in this chapter, I expore how my system can 
perform extended applied ethical reasoning. Metaethical reasoning is most interesting to philosophers 
who are trying to formulate the ``best" theory of ethics. Applied ethical reasoning, on the other hand, 
is useful to ordinary people who are trying to decide how to live their lives. In order for automated 
ethics to guide AI agents, it must perform applied ethical reasoning.

One challenge of applied ethical reasoning is that it requires more common sense knowledge than metaethical
reasoning. Because metaethics is about ethics itself, and not about the dilemmas that ethics is 
supposed to help us resolve, this kind of reasoning requires very little knowledge about the world. 
In previous chapters, I perform metaethical reasoning using my system (which formalizes an 
ethical theory) and basic logical facts and objects. Applied ethical reasoning, on the other hand,
focuses on a particular ethical dilemma and thus requires enough 
``common sense" to understand the dilemma and options at hand. For example, an applied ethicist 
evaluating the permissibility of lying needs some robust definition of the term lying and likely some
understanding about the activities of communication and truth telling. Kantians specifically describe
this common sense as ``postulates of rationality" that are nontrivial and nonnormative, but still
part of the process of practical reasoning itself \citep{silber}. \citet{powers} notes that common sense 
is a hurder for an automated Kantian ethical agent. In this chapter, I attempt to tackle this challenge and
endow my system with this kind of common sense in the specific case of lying. In order for my 
system to be used to perform applied ethical reasoning, it will need to be equipped with a database
of common sense facts and definitions as I present in this chapter. My system will contribute the core
reasoning about the Formula of Universal Law, but this reasoning must be applied to objects that are defined
in this common sense database. In this chapter, I try to understand what this common sense reasoning looks like
for the specific example of lying.

@@

@@

\isamarkupcmt{Unlike many of the other proofs in this project, this proof was a little heavier and required some manual
work to produce. I divided the proof into the intermediate steps shown here, and Isabelle was able to do 

@@

deception" \citep[4]{KorsgaardRTL}. First, I define a joke..%

@@

Korsgaard argues that her above argument for a prohibition against lying also implies that joking is 
permissible, because its purpose is \emph{not} to decieve, but something else entirely. This means that, 
even armed with the same core convention of lying assumption as above, joking must be permissible. The 
lemma below shows exactly that.%
\ joking{\isacharunderscore}not{\isacharunderscore}prohibited{\isacharcolon}\isanewline
\isamarkupcmt{Initial set-up: $m$ is a well-formed maxim composed of some circumstances, act, and goal.%
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}joke\ m{\isachardoublequoteclose}\isanewline
\isamarkupcmt{$m$ is a maxim about joking. Precisely, it is a maxim in which the action is to knowingly utter a 
falsehood and the goal does not require that someone believe this falsehood.%
\isamarkupcmt{The same convention of lying assumption as in the above example.%
worlds for this example. An irrelevant technical detail.%
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymTurnstile}\ {\isacharparenleft}permissible\ m\ me{\isacharparenright}{\isachardoublequoteclose}\isanewline
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}smt\ assms{\isacharparenleft}{\isadigit{1}}{\isacharparenright}\ assms{\isacharparenleft}{\isadigit{2}}{\isacharparenright}\ assms{\isacharparenleft}{\isadigit{3}}{\isacharparenright}\ case{\isacharunderscore}prod{\isacharunderscore}conv{\isacharparenright}%
This completes my implementation of Korsgaard's first interpretation of lying. Using the convention
of lying assumption formalized in this section, I show that lying is generally prohibited but joking
is permissible. Notice that this is already a sophisticated kind of ethical reasoning. The sophistication
necessary to distinguish between lying and joking is a direct consequence of my system's use of a robust
conception of a maxim, which encodes the goal of an act as part of the maxim being evaluated. This kind 
of ethical reasoning also requires relatively few and uncontroversial common sense facts. The deepest
assumption required was that, if everyone lies about a given statement, no one will believe that 
statement. This assumption is not merely definitional; it does encode some synthetic knowledge about the 
world, but it is relatively uncontroversial. Indeed, it is so well-accepted that Korsgaard does not 
bother to justify it in her argument. These examples demonstrate that, while common sense reasoning 
is an obstacle that must be overcome for my system to be used in practice, it is surmountable.%
\isamarkupsubsection{Lying to a Liar \label{sec:lyingtoliar}%
Once Korsgaard completes her preliminary work differentiating between lies and jokes, she begins her main
 argument, which examines the controversial case of the murderer at the door. Recall that the murderer
appears at your door and asks if his intended victim is at home. Ordinary intuition requires that it 
is at the very least permissible (if not obligatory) to lie to the murderer in order to protect the 
victim. Korsgaard notes that a murderer who wishes to find his victim cannot simply announce his
intentions to murder; instead, he must ``must suppose that you do not know who he is and
what he has in mind" \citep[5]{KorsgaardRTL}.\footnote{Korsgaard assumes that the murderer will lie 
about his identity in order to take advantage of your honesty. In footnote 5, she accepts that her 
arguments will not apply in the case of the honest murderer who announces his intentions, so she 
restricts her focus to the case of lying to a liar. She claims that in the case of the honest murderer, the correct act is 
to refuse to respond, but she does not argue for this in this paper.} Thus, she can modify the maxim in 
question to specify that when someone lies to you, you are allowed to lie to them. The maxim of lying 
to the murderer is actually the maxim of lying to a liar, which she argues is permissible. Notice that
her argument hinges on this clever, but ultimately sensical formulation of your maxim. She notes
that there is something relevant and significant about the fact that the person demanding to know Sara's
location is a murderer and that he is trying to take advantage of your honesty. This claim is not
unfounded or wildly controversial, but it does demonstrate the importance of correctly formulating
the maxim to test. In this section, I will formalize her argument for lying to a liar.

@@

Again, this is an uncontroversial statement; the murderer assumes that if you knew he was a murderer, you 
would not disclose the victim's location to him. SHOULD I SAY MORE HERE?%

@@

wills it. Korsgaard assumes that if everyone wills a maxim, they also believe that everyone else wills it.
DO I NEED TO SAY MORE HERE?%

@@

in order to find a victim is not. The assumptions used in this example are a little more robust, but still
ultimately uncontroversial. Indeed, they are direct consequences of Korsgaard's definition of willing 

@@

that Kantian scholars debate robustly. Indeed, armed with this common sense, my system generated 

@@

While it is true that lying to the murderer should be permissible, Korsgaard notes that many will want
to say something stronger, like the fact that lying to the murderer is obligatory in order to protect
the intended victim \citep[15]{KorsgaardRTL}. It seems like we would be doing something wrong if I revealed the victim's
location, knowing that this revelation would cost them their life. Korsgaard solves this problem by 
noting that, while the FUL shows that lying to the murderer permissible, other parts of Kant's ethics
will show that it is obligatory. Recall that Kant presents perfect and imperfect duties,
where the former are strict, inviolable, and specific and the latter are broader prescriptions for action.
Perfect duties always supersede imperfect duties when the two conflict.
For example, the duty to not murder is a perfect duty and the duty to give to charity is an imperfect duty. 
The FUL generates perfect duties and Kant's extended theory of virtues generates imperfect duties. The details
of this theory and these distinctions are outside the scope of this paper, but the crucial note is that 
other parts of Kant's ethical theory generate the obligation to lie to the murderer. I chose
to formalize the FUL because it is, in some sense, the strongest of the formulations of the categorical
imperative. An even more sophisticated Kantian reasoner could formalize his theory of virtue and his
other formulatations of the categorical imperative in order to generate the obligation to lie to the murderer, 
but the FUL is the strongest and most foundational of these principles. The fact that my system merely
shows that lying to the murderer is permissible, but not obligatory is consistent with the part of 
Kant's ethical theory that I formalize and demonstrates that I have faithfully implemented the FUL.%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
While this example demonstrates the power of my system (when equipped with some common sense), it 
also shows how vital the role of the common sense reasoning is. Slight, intuitive changes in the common
sense facts can achieve totally different conclusions about lying. This represents an obstacle to 
fully automated ethical reasoning; such an agent would need a trusted database of common 
sense facts, which is still an unsolved problem. My work is one step towards such an agent, but the 
importance of common sense means that much progress must be made in order to completely automate ethics.

The reasoning of this section also demonstrates one additional place where a Kantian must make vital judgements:
the formulation of the maxim itself. Korsgaard's argument for the permissibility of lying to a 
murderer hinged on a clever formulation of the maxim highlighting a particular facet of the circumstances, 
namely that the murderer is lying to you. Indeed, there is robust debate in the literature on what
circumstances should be considered when formulating a maxim. Some critics of Kant raise the ``tailoring
objection," which is the worry that arbitrarily specific circumstances render any maxim universalizable. For 
example, the maxim ``When my name is Lavanya Singh and I am wearing a purple shirt and it is November 26th, 
I will lie in order to get some easy cash" passes the universalizability test. Even if this maxim is 
willed universally, the circumstances are so specific that lying will not become the general mechanism
for getting easy cash, so the lender will believe my lie and the maxim will remain effective. By tailoring
the circumstances, any maxim can evade universalization.

The Kantian response to this criticism is to require that the circumstances included in the formulation
of the maxim be ``morally relevant." In the example above, my purple shirt and the date clearly have no bearing on 
the moral status of lying. On the other hand, consider the maxim ``When I am unmeployed, I will murder
someone in order to take their job." The circumstances of being unmeployed clearly have some bearing on the moral
relevance of the murder in question; they speak to the motivation for the murder. While this view seems 
to track how we actually perform moral reasoning, it leaves open the question of how to determine
which circumstances are morally relevant. Here, O'Niell reminds us that the Formula of Universal Law is 
a ``test of moral worth rather than of outward rightness" \citep[98]{constofreason}. The FUL is a way 
for an agent to decide how they should behave, not for a third-party to judge their behavior. Ethics is 
a personal process for Kant, so the FUL is designed to help agents internally make decisions, not to 
judge others' decisions. Because agents use the FUL to evaluate their own behavior, the test is at its 
best when they make a good faith effort to isolate the \emph{principle} of their action, rather than some
``surface intent" \citep[87]{constofreason}. The FUL is supposed to determine if an agent's principle of action
is universally consistent, so it is at its most effective when an agent accurately formulates the principle
they act on. Circumstances are morally relevant if they accurately reflect the way that the agent is 
thinking about their own action. In the example above, the circumstances of wearing a purple shirt don't
the principle of the liar's action. They are clearly a disingenous attempt to evade the universalizability
test, but because the FUL is a test of personal integrity, it cannot withstand this kind of mental
gymnastics.

@@

of my system in practice. In this chapter, I argue that using manual, human involvement, these challenges

@@

%:%129=94%:%
%:%141=96%:%
%:%142=97%:%
%:%143=98%:%
%:%144=99%:%
%:%145=100%:%
%:%146=101%:%
%:%147=102%:%
%:%148=103%:%
%:%149=104%:%

@@

@@

%:%160=114%:%
%:%162=115%:%
%:%163=116%:%
%:%164=117%:%
%:%165=118%:%
%:%166=119%:%
%:%167=120%:%
%:%168=121%:%
%:%169=121%:%
%:%170=122%:%
%:%171=123%:%
%:%172=123%:%
%:%174=124%:%
%:%175=124%:%
%:%176=125%:%
%:%177=125%:%
%:%178=126%:%
%:%180=127%:%
%:%181=127%:%
%:%182=128%:%
%:%183=128%:%
%:%184=129%:%
%:%186=130%:%
%:%187=131%:%
%:%188=132%:%
%:%189=132%:%
%:%190=137%:%
%:%191=138%:%
%:%192=138%:%
%:%193=139%:%
%:%195=140%:%
%:%196=141%:%
%:%197=142%:%
%:%200=144%:%
%:%201=145%:%
%:%202=146%:%
%:%203=147%:%
%:%204=148%:%
%:%205=149%:%
%:%206=150%:%
%:%210=152%:%
%:%211=153%:%
%:%213=155%:%
%:%214=155%:%
%:%216=156%:%
%:%217=156%:%
%:%218=157%:%
%:%219=157%:%
%:%221=158%:%
%:%222=159%:%
%:%223=159%:%
%:%224=160%:%
%:%225=160%:%
%:%227=161%:%
%:%230=163%:%
%:%231=164%:%
%:%232=165%:%
%:%233=166%:%
%:%235=168%:%
%:%236=168%:%
%:%237=169%:%
%:%238=170%:%
%:%240=171%:%
%:%241=171%:%
%:%242=172%:%
%:%244=173%:%
%:%245=174%:%
%:%246=174%:%
%:%247=175%:%
%:%249=176%:%
%:%250=177%:%
%:%251=178%:%
%:%252=179%:%
%:%253=180%:%
%:%254=181%:%
%:%255=181%:%
%:%256=182%:%
%:%258=183%:%
%:%259=184%:%
%:%260=184%:%
%:%261=185%:%
%:%268=186%:%
%:%269=186%:%
%:%270=187%:%
%:%271=187%:%
%:%272=188%:%
%:%273=188%:%
%:%275=189%:%
%:%276=190%:%
%:%277=191%:%
%:%278=192%:%
%:%279=192%:%
%:%280=193%:%
%:%281=193%:%
%:%282=194%:%
%:%283=194%:%
%:%284=195%:%
%:%285=195%:%
%:%286=196%:%
%:%287=196%:%
%:%288=196%:%
%:%290=197%:%
%:%291=198%:%
%:%292=198%:%
%:%293=199%:%
%:%303=201%:%
%:%304=202%:%
%:%305=203%:%
%:%306=204%:%
%:%307=205%:%
%:%309=207%:%
%:%310=207%:%
%:%311=208%:%
%:%313=210%:%
%:%314=211%:%
%:%315=212%:%
%:%316=213%:%
%:%317=214%:%
%:%318=215%:%

@@

%:%321=218%:%

@@

@@

%:%325=221%:%
%:%326=222%:%

>>>>

%
\begin{isabellebody}%
\setisabellecontext{application}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isatagtheory
%
\endisatagtheory
{\isafoldtheory}%
%
\isadelimtheory
%
\endisadelimtheory
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsection{Applications \label{sec:applications}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
In this chapter, I 
demonstrate my system's ability to formalize longer, more complicated ethical arguments and present 
the additional capabilities necessary to use my system in practice. In Chapter 2
and Chapter 3, I performed metaethical reasoning while testing different formalizations of the FUL. Metaethical
reasoning analyzes properties of moral thought itself and involves questions about the nature of ethical
truth. This kind of reasoning contrasts with ``applied 
ethical reasoning," which is the use of ethics to resolve dilemmas and make judgements about 
what an agent should or should not do. Applied ethical reasoning is relevant with respect to an agent's
particular situation, not just to an ethical theory as an abstract entity. In previous chapters' application tests,
I performed some toy examples of applied ethical reasoning, but in this chapter, I expore how my system can 
perform extended applied ethical reasoning. Metaethical reasoning is most interesting to philosophers 
who are trying to formulate the ``best" theory of ethics. Applied ethical reasoning, on the other hand, 
is useful to ordinary people who are trying to decide how to live their lives. In order for automated 
ethics to guide AI agents, it must perform applied ethical reasoning.

One challenge of applied ethical reasoning is that it requires more common sense knowledge than metaethical
reasoning. Because metaethics is about ethics itself, and not about the dilemmas that ethics is 
supposed to help us resolve, this kind of reasoning requires very little knowledge about the world. 
In previous chapters, I perform metaethical reasoning using my system (which formalizes an 
ethical theory) and basic logical facts and objects. Applied ethical reasoning, on the other hand,
focuses on a particular ethical dilemma and thus requires enough 
``common sense" to understand the dilemma and options at hand. For example, an applied ethicist 
evaluating the permissibility of lying needs some robust definition of the term lying and likely some
understanding about the activities of communication and truth telling. Kantians specifically describe
this common sense as ``postulates of rationality" that are nontrivial and nonnormative, but still
part of the process of practical reasoning itself \citep{silber}. \citet{powers} notes that common sense 
is a hurdle for an automated Kantian ethical agent. In this chapter, I attempt to tackle this challenge and
endow my system with this kind of common sense in the specific case of lying. In order for my 
system to be used to perform applied ethical reasoning, it will need to be equipped with a database
of common sense facts and definitions as I present in this chapter. My system will contribute the core
reasoning about the Formula of Universal Law, but this reasoning must be applied to objects that are defined
in this common sense database. In this chapter, I try to understand what this common sense reasoning looks like
for the specific example of lying.

@@

In this paper, I contribute an implementation of the Formula of Universal Law that, when equipped with 
relatively thin common sense facts, can performed nuanced, sophisticated ethical reasoning. When given 
as input a maxim in the format specified in Chapter 3, my system can show that the maxim is permissible, 
obligatory, or prohibited. It can provide a verifiable proof of its judgement and a human-readable list
of the facts or axioms it used to reach its conclusion. My system is faithful to philosophical literatire
on Kantian ethics and serves as the first step towards a fully or partially automated ethical AI agent. 
It also establishes the power and potential of computational ethics, or the use of computational tools
for ethical investigation, as explored in this section's Applied Ethical reasoning.

@@

\isamarkupcmt{Unlike many of the other proofs in this project, this proof is a little heavier and requires some manual
work to produce. After I divide the proof into the intermediate steps shown here, and Isabelle is able to do 

@@

deception" \citep[4]{KorsgaardRTL}. First, I define a joke.%

@@

Korsgaard argues that her above argument for a prohibition against lying also implies that joking is 
permissible, because its purpose is \emph{not} to decieve, but something else entirely. This means that, 
even armed with the same core convention of lying assumption as above, joking must be permissible. The 
lemma below shows exactly that.%
\end{isamarkuptext}\isamarkuptrue%
\isacommand{lemma}\isamarkupfalse%
\ joking{\isacharunderscore}not{\isacharunderscore}prohibited{\isacharcolon}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}m\ {\isasymequiv}\ {\isacharparenleft}c{\isacharcolon}{\isacharcolon}t{\isacharcomma}\ a{\isacharcolon}{\isacharcolon}os{\isacharcomma}\ g{\isacharcolon}{\isacharcolon}t{\isacharparenright}{\isachardoublequoteclose}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymforall}w{\isachardot}\ {\isasymforall}s{\isachardot}\ well{\isacharunderscore}formed\ m\ s\ w{\isachardoublequoteclose}\isanewline
%
\isamarkupcmt{Initial set-up: $m$ is a well-formed maxim composed of some circumstances, act, and goal.%
}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}joke\ m{\isachardoublequoteclose}\isanewline
%
\isamarkupcmt{$m$ is a maxim about joking. Precisely, it is a maxim in which the action is to knowingly utter a 
falsehood and the goal does not require that someone believe this falsehood.%
}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymforall}t\ w{\isachardot}\ {\isacharparenleft}{\isacharparenleft}{\isasymforall}p{\isachardot}\ utter{\isacharunderscore}falsehood\ p\ t\ w{\isacharparenright}\ {\isasymlongrightarrow}\ {\isacharparenleft}{\isasymforall}p{\isachardot}\ \isactrlbold {\isasymnot}\ {\isacharparenleft}believe\ p\ t{\isacharparenright}\ w{\isacharparenright}{\isacharparenright}{\isachardoublequoteclose}\isanewline
%
\isamarkupcmt{The same convention of lying assumption as in the above example.%
}\isanewline
\ \ \isakeyword{assumes}\ {\isachardoublequoteopen}{\isasymforall}w{\isachardot}\ c\ w{\isachardoublequoteclose}\isanewline
%
\isamarkupcmt{Restrict our focus to worlds in which the circumstances hold, as these are the morally interesting 
worlds for this example. An irrelevant technical detail.%
}\isanewline
\ \ \isakeyword{shows}\ {\isachardoublequoteopen}{\isasymTurnstile}\ {\isacharparenleft}permissible\ m\ me{\isacharparenright}{\isachardoublequoteclose}\isanewline
%
\isadelimproof
\ \ %
\endisadelimproof
%
\isatagproof
\isacommand{by}\isamarkupfalse%
\ {\isacharparenleft}smt\ assms{\isacharparenleft}{\isadigit{1}}{\isacharparenright}\ assms{\isacharparenleft}{\isadigit{2}}{\isacharparenright}\ assms{\isacharparenleft}{\isadigit{3}}{\isacharparenright}\ case{\isacharunderscore}prod{\isacharunderscore}conv{\isacharparenright}%
\endisatagproof
{\isafoldproof}%
%
\isadelimproof
%
\endisadelimproof
%
\begin{isamarkuptext}%
One potential worry with the above argument is the fact that my definition of joke requires that 
achieving the goal of a joke does not rely on anyone believing the falsehood told in the joke. On its face, 
this is not a complete representation of a joke; the ordinary conception of a joke would be something like 
a false statement uttered with the goal of amusing. I define the goal of the joke as not requiring that 
anyone believe the false statement to distinguish it from lying, which has deception as the goal. I 
intentionally avoid providing robust formalizations of both deception and amusement; instead I present a 
thin conception that relies entirely on the concept of belief. This conception does not require any notion
of humor, satire, intention, or malice, all of which are concepts that may be necessary to define 
amusement and deception completely.

The fact that my system shows that lies are prohibited and jokes are permissible with these things 
conceptions of amusement and deception shows that my system isolates a necessary and sufficient property 
of a class of maxims that fail the universalizability test. Because my definitions of a lie and joke
only differ in whether or not their goal requires that someone believe the falsehood in question, the
theorems proved in this section show that, in order for a maxim with the act of knowingly uttering a falsehood
to be prohobited, the goal requires that someone believe the falsehood is a necessary and sufficient condition. 
This logical fact derived by my system tracks a fact implicit in Korsgaard's argument and in most Kantian accounts
of lying: the wrongness of lying is derived from the requirement that someone believe the falsehood. 
The logical reality that this property is necessary and sufficient to generate a prohibition reflects
a deep philosohopical explanation of \emph{why} certain maxims about uttering falsehood fail the universalizability
test and why others pass. In simple terms, universalizing uttering a falsehood makes belief in that 
falsehood impossible, so any maxims with goals that require believing in the falsehood will be prohibited.

This account not only describes the kind of maxims that fail or pass the universalizability test, it 
also provides a guide to constructing permissible maxims about uttering falsehoods. As an example, 
consider the idea of throwing a surprise birthday party. At first glance, the maxim of action is 
something like, ``When it is my friend's birthday, I will secretly plan a party so that I can surprise
them." The goal ``so that I can surprise them" clearly requires that your friend believe the falsehood that 
you are not planning a party, else the surprise would be ruined. The analysis above seems to imply that 
Kantian ethics would prohibit surprise parties, which is a sad conclusion for birthday-lovers everywhere. 
Noticing that the problem with this maxim is that the goal requires belief
in a falsehood provides a way to rescue the beloved concept of surprise parties. When throwing a 
surprise party, the ultimate objective is \emph{not} to surprise your friend, but is to celebrate
your friend and help them have a fun birthday. If someone ruins the surprise, but the party is still fun
and the birthday person still feels loved, then we would consider such a party a success! Someone who
called this party a failure clearly would be missing the point of a surprise party. Thus, the goal of 
a surprise party is not the surprise itself, but rather to celebrate the birthday person. Thus modified, 
the goal no longer requires belief in the falsehood and thus passes the universalizability test.

The implications of this section are twofold. First, my system is capable of performing ethical reasoning
sophisticated enough to show that lying is prohibited but joking is not. The sophistication
necessary to distinguish between lying and joking was a direct consequence of my system's use of a robust
conception of a maxim, which encoded the goal of an act as part of the maxim being evaluated. Second, 
in the process of making this argument precise, my system isolated a necessary and sufficient condition 
of a maxim about uttering a falsehood being prohibited: that the goal require that someone believe
the falsehood. This condition both made an long-standing argument in Kantian ethics more precise
and can guide the correct formulation of future maxims. In other words, an insight generated by the 
computer provides value to ethicists.

Moreover, all of the reasoning in this chapter required relatively few and uncontroversial common sense facts. The deepest
assumption required was that, if everyone lies about a given statement, no one will believe that 
statement. This assumption is not merely definitional; it does encode some synthetic knowledge about the 
world, but it is relatively uncontroversial. Indeed, it is so well-accepted that Korsgaard does not 
bother to justify it in her argument. These examples showed that, while common sense reasoning 
is an obstacle that must be overcome for my system to be used in practice, it is surmountable.%
\end{isamarkuptext}\isamarkuptrue%
%
\isadelimdocument
%
\endisadelimdocument
%
\isatagdocument
%
\isamarkupsubsection{Lying to a Liar \label{sec:lyingtoliar}%
}
\isamarkuptrue%
%
\endisatagdocument
{\isafolddocument}%
%
\isadelimdocument
%
\endisadelimdocument
%
\begin{isamarkuptext}%
Once Korsgaard completes her preliminary work differentiating between lies and jokes, she begins her main
 argument, which examines the controversial case of the murderer at the door. Recall that the murderer
appears at your door and asks if his intended victim is at home. Ordinary intuition requires that it 
is at the very least permissible (if not obligatory) to lie to the murderer in order to protect the 
victim. Korsgaard notes that a murderer who wishes to find his victim cannot simply announce his
intentions to murder; instead, he must ``must suppose that you do not know who he is and
what he has in mind" \citep[5]{KorsgaardRTL}.\footnote{Korsgaard assumes that the murderer will lie 
about his identity in order to take advantage of your honesty. In footnote 5, she accepts that her 
arguments will not apply in the case of the honest murderer who announces his intentions, so she 
restricts her focus to the case of lying to a liar. She claims that in the case of the honest murderer, the correct act is 
to refuse to respond, but she does not argue for this in this paper.} Thus, she can modify the maxim in 
question to specify that when someone lies to you, you are allowed to lie to them. The maxim of lying 
to the murderer is actually the maxim of lying to a liar, which she argues is permissible. Notice that
her argument hinges on this clever, but ultimately sensical formulation of your maxim. She notes
that there is something relevant and significant about the fact that the person demanding to know Sara's
location is a murderer and that he is trying to take advantage of your honesty. This claim is not
unfounded or wildly controversial, but it does demonstrate the importance of correctly formulating
the maxim to test. In this section, I will formalize her argument for lying to a liar.

@@

Again, this is Korsgaard's uncontroversial assumption; the murderer assumes that if you knew he was a murderer, you 
would not disclose the victim's location to him.%

@@

wills it and thus notices the pattern of everyone willing it. If you observe that many do X in circumstances C,
you will assume that everyone does X in circumstance C.%

@@

in order to find a victim is not. The assumptions used in this example were a little more robust, but still
ultimately uncontroversial because they were direct consequences of Korsgaard's definition of willing 

@@

that Kantian scholars debate robustly. Armed with this common sense, my system generated 

@@

While it is true that lying to the murderer should be permissible, Korsgaard notes that many will want
to say something stronger, like the fact that lying to the murderer is obligatory in order to protect
the intended victim \citep[15]{KorsgaardRTL}. It seems like we would be doing something wrong if I revealed the victim's
location, knowing that this revelation would cost them their life. Korsgaard solves this problem by 
noting that, while the FUL shows that lying to the murderer permissible, other parts of Kant's ethics
show that it is obligatory. Recall that Kant presents perfect and imperfect duties,
where the former are strict, inviolable, and specific and the latter are broader prescriptions for action.
Perfect duties always supersede imperfect duties when the two conflict.
For example, the duty to not murder is a perfect duty and the duty to give to charity is an imperfect duty. 
The FUL generates perfect duties and Kant's extended theory of virtues generates imperfect duties. The details
of this theory and these distinctions are outside the scope of this paper, but the crucial note is that 
other parts of Kant's ethical theory generate the obligation to lie to the murderer. I chose
to formalize the FUL because it is, in some sense, the strongest of the formulations of the categorical
imperative. An even more sophisticated Kantian reasoner could formalize his theory of virtue and his
other formulatations of the categorical imperative in order to generate the obligation to lie to the murderer, 
but the FUL is the strongest and most foundational of these principles. The fact that my system merely
shows that lying to the murderer is permissible, but not obligatory is consistent with the part of 
Kant's ethical theory that I formalize and demonstrates that I have faithfully implemented the FUL.%
\end{isamarkuptext}\isamarkuptrue%
%
\begin{isamarkuptext}%
While this example demonstrates the power of my system (when equipped with some common sense), it 
also shows how vital the role of the common sense reasoning is. Slight, intuitive changes in the common
sense facts achieved totally different conclusions about lying. This represents an obstacle to 
fully automated ethical reasoning; such an agent would need a trusted database of common 
sense facts, which is still an unsolved problem. My work is one step towards such an agent, but the 
importance of common sense means that much progress must be made in order to completely automate ethics.

The reasoning of this section also demonstrated one additional place where a Kantian must make vital judgements:
the formulation of the maxim itself. Korsgaard's argument for the permissibility of lying to a 
murderer hinged on a clever formulation of the maxim highlighting a particular facet of the circumstances, 
namely that the murderer is lying to you. Indeed, there is robust debate in the literature on what
circumstances should be considered when formulating a maxim. Some critics of Kant raise the ``tailoring
objection," which is the worry that arbitrarily specific circumstances render any maxim universalizable. For 
example, the maxim ``When my name is Lavanya Singh and I am wearing a purple shirt and it is November 26th, 
I will lie in order to get some easy cash" passes the universalizability test. Even if this maxim is 
willed universally, the circumstances are so specific that lying will not become the general mechanism
for getting easy cash, so the lender will believe my lie and the maxim will remain effective. By tailoring
the circumstances, any maxim can evade universalization.

The Kantian response to this criticism is to require that the circumstances included in the formulation
of the maxim be ``morally relevant." In the example above, my purple shirt and the date clearly have no bearing on 
the moral status of lying. On the other hand, consider the maxim, ``When I am unmeployed, I will murder
someone in order to take their job." The circumstances of being unmeployed clearly have some bearing on the moral
relevance of the murder in question; they speak to the motivation for the murder. While this view seems 
to track how we actually perform moral reasoning, it leaves open the question of how to determine
which circumstances are morally relevant. Here, O'Niell reminds us that the Formula of Universal Law is 
a ``test of moral worth rather than of outward rightness" \citep[98]{constofreason}. The FUL is a way 
for an agent to decide how they should behave, not for a third-party to judge their behavior. Ethics is 
a personal process for Kant, so the FUL is designed to help agents internally make decisions, not to 
judge others' decisions. Because agents use the FUL to evaluate their own behavior, the test is at its 
best when they make a good faith effort to isolate the \emph{principle} of their action, rather than some
``surface intent" \citep[87]{constofreason}. The FUL is supposed to determine if an agent's principle of action
is universally consistent, so it is at its most effective when an agent accurately formulates the principle
they act on. Circumstances are morally relevant if they accurately reflect the way that the agent is 
thinking about their own action. In the example above, the circumstances of wearing a purple shirt don't reflect
the principle of the liar's action. Their inclusion is clearly a disingenous attempt to evade the universalizability
test, but because the FUL is a test of personal integrity, it cannot withstand this kind of mental
gymnastics.

@@

of my system in practice. In this chapter, I argued that using manual, human involvement, these challenges

@@

%:%121=93%:%
%:%122=94%:%
%:%123=95%:%
%:%124=96%:%
%:%125=97%:%
%:%126=98%:%
%:%127=99%:%
%:%128=100%:%
%:%129=101%:%
%:%138=103%:%

@@

%:%158=113%:%

@@

%:%160=115%:%
%:%161=116%:%
%:%162=117%:%
%:%163=118%:%
%:%164=119%:%
%:%165=120%:%
%:%166=121%:%
%:%168=123%:%
%:%169=123%:%
%:%171=124%:%
%:%172=125%:%
%:%173=126%:%
%:%174=127%:%
%:%175=128%:%
%:%176=129%:%
%:%177=130%:%
%:%178=130%:%
%:%179=131%:%
%:%180=132%:%
%:%181=132%:%
%:%183=133%:%
%:%184=133%:%
%:%185=134%:%
%:%186=134%:%
%:%187=135%:%
%:%189=136%:%
%:%190=136%:%
%:%191=137%:%
%:%192=137%:%
%:%193=138%:%
%:%195=139%:%
%:%196=140%:%
%:%197=141%:%
%:%198=141%:%
%:%199=146%:%
%:%200=147%:%
%:%201=147%:%
%:%202=148%:%
%:%204=149%:%
%:%205=150%:%
%:%206=151%:%
%:%209=153%:%
%:%210=154%:%
%:%211=155%:%
%:%212=156%:%
%:%213=157%:%
%:%214=158%:%
%:%215=159%:%
%:%219=161%:%
%:%220=162%:%
%:%222=164%:%
%:%223=164%:%
%:%225=165%:%
%:%226=165%:%
%:%227=166%:%
%:%228=166%:%
%:%230=167%:%
%:%231=168%:%
%:%232=168%:%
%:%233=169%:%
%:%234=169%:%
%:%236=170%:%
%:%239=172%:%
%:%240=173%:%
%:%241=174%:%
%:%242=175%:%
%:%244=177%:%
%:%245=177%:%
%:%246=178%:%
%:%247=179%:%
%:%249=180%:%
%:%250=180%:%
%:%251=181%:%
%:%253=182%:%
%:%254=183%:%
%:%255=183%:%
%:%256=184%:%
%:%258=185%:%
%:%259=186%:%
%:%260=187%:%
%:%261=188%:%
%:%262=189%:%
%:%263=190%:%
%:%264=190%:%
%:%265=191%:%
%:%267=192%:%
%:%268=193%:%
%:%269=193%:%
%:%270=194%:%
%:%277=195%:%
%:%278=195%:%
%:%279=196%:%
%:%280=196%:%
%:%281=197%:%
%:%282=197%:%
%:%284=198%:%
%:%285=199%:%
%:%286=200%:%
%:%287=201%:%
%:%288=201%:%
%:%289=202%:%
%:%290=202%:%
%:%291=203%:%
%:%292=203%:%
%:%293=204%:%
%:%294=204%:%
%:%295=205%:%
%:%296=205%:%
%:%297=205%:%
%:%299=206%:%
%:%300=207%:%
%:%301=207%:%
%:%302=208%:%
%:%312=210%:%
%:%313=211%:%
%:%314=212%:%
%:%315=213%:%
%:%316=214%:%
%:%318=216%:%

@@

@@

%:%323=220%:%

@@

%:%325=222%:%
%:%326=223%:%
%:%327=224%:%
